Pteropod physiological analysis

### 2018-04-12 physio.Rmd
only mods with low VIF, only select response/exp vars, model sel results, removed methods text

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)

@@

# Physiological correlations with biomarkers {.tabset}

```{r}
tochk <- ptedat %>% 
  select(-CTD, -Latitude)

# all correlations
crs <- crossing(var1 = names(tochk), var2 = names(tochk)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- tochk[[x$var1]]
      vr2 <- tochk[[x$var2]]

@@

      # spearman
      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
      sp_cr <- round(sp_ts$estimate, 2)
      sp_pv <- p_ast(sp_ts$p.value)
      sp <- paste(sp_cr, sp_pv)

      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)

sp_cr <- crs %>% 
  select(var1, var2, sp) %>% 
  spread(var2, sp) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1')

pr_cr <- crs %>% 
  select(var1, var2, pr) %>% 
  spread(var2, pr) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1') 
```

## Spearman correlations

```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

```{r , fig.height = 5, fig.width = 10}
spplo <- crs %>% 
  select(-pr) %>% 
  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations
<div class="verysmall">

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

```{r , fig.height = 5, fig.width = 10}
prplo <- crs %>% 
  select(-sp) %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

# Physiological correlations with environment {.tabset}

```{r}
# all correlations
crs <- crossing(var1 = names(dat_frm), var2 = names(dat_frm)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- dat_frm[[x$var1]]
      vr2 <- dat_frm[[x$var2]]

      # spearman
      sp_ts <- suppressWarnings(cor.test(vr1, vr2, method = 'spearman'))
      sp_cr <- round(sp_ts$estimate, 2)
      sp_pv <- p_ast(sp_ts$p.value)
      sp <- paste(sp_cr, sp_pv)

      out <- data.frame(pr = pr, sp = sp, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)

sp_cr <- crs %>% 
  select(var1, var2, sp) %>% 
  spread(var2, sp) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1')

pr_cr <- crs %>% 
  select(var1, var2, pr) %>% 
  spread(var2, pr) %>% 
  data.frame(stringsAsFactors = F) %>% 
  remove_rownames %>% 
  column_to_rownames('var1') 
```

## Spearman correlations
<div class="verysmall">

```{r}
knitr::kable(sp_cr, caption = 'Spearman rank correlations')
```

```{r spheat, fig.height = 5, fig.width = 10}
spplo <- crs %>% 
  select(-pr) %>% 
  separate(sp, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(spplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

## Pearson correlations
<div class="verysmall">

```{r}
knitr::kable(pr_cr, caption = 'Pearson correlations')
```

```{r prheat, fig.height = 5, fig.width = 10}
prplo <- crs %>% 
  select(-sp) %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(unique(.$var1)))
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = 'top', 
  plot.margin = grid::unit(c(2, 2, 2, 2), 'pt'), 
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

p <- ggplot(prplo, aes(y = var1, x = var2, fill = cor)) + 
  geom_tile(colour = 'black') + 
  pbase +
  scale_y_discrete('', expand = c(0, 0)) + 
  scale_x_discrete('', expand = c(0, 0)) +
  scale_fill_gradient2('Correlation', low = muted("green"), mid = "white", high = muted("blue"), midpoint = 0) +
  guides(fill = guide_colourbar(barheight = 0.5, barwidth = 5, label.theme = element_text(size = 6, angle = 0)))
p
```

@@

@@

Pairwise models with interactions for each biomarker, only pairwise models with VIF < 11 are shown

```{r, fig.height = 9}
# model data
mods <- phymod %>% 
  unnest(data)
mod_all <- phymod

@@

## type 2 dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ty2'), caption = 'ty2')
```

## type 3 dissolution
```{r}
knitr::kable(filter(mods, pte_lab %in% 'ty3'), caption = 'ty3')
```

## scarring
```{r}
knitr::kable(filter(mods, pte_lab %in% 'scr'), caption = 'scr')
```

## length
```{r}
knitr::kable(filter(mods, pte_lab %in% 'len'), caption = 'len')
```

>>>>

library(tidyverse)
library(vegan)
library(ggord)
library(scales)
library(effects)
library(shiny)
library(plotly)
library(grid)

@@

# Correlations

```{r}
# physiology rda model
dat_cor <- ptedat %>% 
  select(one_of('CTD', ptechr)) %>% 
  # na.omit %>% 
  left_join(envdat, by = 'CTD') %>% 
  data.frame %>% 
  remove_rownames %>% 
  column_to_rownames('CTD')

# all correlations
crs <- crossing(var1 = names(dat_cor), var2 = names(dat_cor)) %>% 
  filter(var1 != var2) %>% 
  rownames_to_column() %>% 
  group_by(rowname) %>% 
  nest %>% 
  mutate(
    crs = map(data, function(x){

      # variables
      vr1 <- dat_cor[[x$var1]]
      vr2 <- dat_cor[[x$var2]]

@@

      out <- data.frame(pr = pr, stringsAsFactors = F)
      return(out)

    })
  ) %>% 
  unnest %>% 
  select(-rowname)
```

```{r, fig.height = 6, fig.width = 6}
levs <- c(sort(envchr), sort(ptechr))
labs <- c('Omega[ar]', 'chla', 'O[2]', 'pCO[2]', 'pH', 'Temp', 'abundance', 'dissolution', 'growth', 'scarring', 'typeII', 'typeIII')
prplo <- crs %>% 
  separate(pr, c('cor', 'sig'), sep = ' ') %>% 
  filter(var1 %in% levs & var2 %in% levs) %>%  
  mutate(
    cor = as.numeric(cor), 
    var1 = factor(var1, levels = rev(levs), labels = rev(labs)), 
    var2 = factor(var2, levels = rev(levs), labels = rev(labs)), 
    sig = gsub('ns', '', sig)
  )

pbase <- theme(
  panel.grid.major = element_blank(), 
  panel.grid.minor = element_blank(), 
  axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8), 
  axis.text.y = element_text(size = 8),
  legend.position = c(0.5, 1.2),
  legend.direction = 'horizontal',
  plot.margin = unit(c(8,4,0,0), "lines"),
  strip.background = element_blank(), 
  strip.text.y = element_text(angle = 0, hjust = 0, vjust = 0.5), 
  panel.background = element_rect(fill = 'black')
  )

outlab <- data.frame(
  y = c(3.5, 9.5), 
  lab = c('Population/\nPhysiology', 'Environment')
)

p <- ggplot(prplo) + 
  geom_tile(aes(y = var1, x = var2, fill = cor), colour = 'black') + 
  geom_text(aes(y = var1, x = var2, label = sig)) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[1], ymax = outlab$y[1], xmin = 13, xmax = 13) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0, gp = gpar(cex = 0.7)),
                    ymin = outlab$y[2], ymax = outlab$y[2], xmin = 13, xmax = 13) +
  annotation_custom(grob = textGrob(label = outlab$lab[1], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[1], xmax = outlab$y[1], ymin = 13.5, ymax = 13.5) +
  annotation_custom(grob = textGrob(label = outlab$lab[2], hjust = 0.5, gp = gpar(cex = 0.7)),
                    xmin = outlab$y[2], xmax = outlab$y[2], ymin = 13.5, ymax = 13.5) +
  pbase +
  scale_y_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) + 
  scale_x_discrete('', expand = c(0, 0), labels = parse(text = rev(labs))) +
  scale_fill_gradient2('Correlation', low = muted("blue"), mid = "white", high = muted("red"), midpoint = 0) +
  geom_hline(yintercept = 6.5, size = 1.5) +
  geom_vline(xintercept = 6.5, size = 1.5)

# Code to override clipping
gt <- ggplot_gtable(ggplot_build(p))
gt$layout$clip[gt$layout$name == "panel"] <- "off"
grid.draw(gt)
```

@@

# Model results {.tabset}

```{r}
# model data
mods <- phymod %>% 
  unnest(data)
mod_all <- phymod
```

@@

Pairwise models with interactions for relevant response measures, only pairwise models with VIF < 11 are shown

```{r, fig.height = 9}

@@
